---
title: "Charlie_16S"
output: html_document
date: "2023-03-03"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

setwd("~/Dropbox/Charlie_16S/scripts/")

knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(wesanderson)
library(ggthemes)
library(reshape2)
library(cowplot)
library(biomformat)

```

#OLD pipeline
OTU_reads <- read_biom("../processed_data/feature-table_Silva.biom")
metadata <- read.csv("../ref/metadata.csv")

zotu <- as.data.frame(as.matrix(biom_data(OTU_reads))) %>%
  cbind(., OTU=paste('OTU',1:length(.$G0046), sep = '')) #%>% #randomly number OTUs
  melt() %>% 
  transform(sample = as.character(variable)) %>%
  select(-variable) %>% rename('value'='counts') %>%
  left_join(., metadata) %>% group_by(sample) %>% 
  mutate(relabd = counts/sum(counts), log10relabd = log10(relabd)) %>%
  mutate(log10relabd = if_else(relabd == 0, -6, log10(relabd))) %>% ungroup()

#identify OTUs with  > 0.5% abundance in at least one sample
otus.to.keep <- zotu %>%
  group_by(OTU) %>%
  summarise(max_abd = max(relabd)) %>%
  filter(max_abd > .001) %>%
  select(OTU) %>% ungroup() %>%
  as_vector() 

#drop samples with under 10K reads...
# Pull out samples w/ >10k reads
samples.to.keep <- zotu %>%
  group_by(sample) %>% mutate(numreads = sum(counts)) %>% 
  ungroup() %>%
  filter(numreads > 10000) %>%
  select(sample) %>% distinct() %>%
  as_vector()
 
# Drop OTUs not at least 0.5% abd in any sample or samples under 10k
zotu <- zotu %>%
  filter(OTU %in% otus.to.keep) %>%
  filter(sample %in% samples.to.keep)

#get taxonomy
tax_table <- as.data.frame(as.matrix(biom_data(OTU_reads))) %>%
  cbind(., OTU=paste('OTU',1:length(.$G0046), sep = '')) %>% mutate(tax_info = row.names(.)) %>% select(OTU, tax_info) %>% 
  separate(tax_info, into = c('x','Kingdom', 'Phylum', 'Class', 'Order', 'Family', 'Genus'), sep = '__') %>% 
        mutate(Kingdom = gsub("\\;[alpha]*","",.$Kingdom),
               Phylum = gsub(";c","",.$Phylum),
               Class = gsub(";o","",.$Class),
               Order = gsub(";f","",.$Order),
               Family = gsub(";g","",.$Family),
               Genus = gsub(";","",.$Genus)) %>% select(-x)

#remove those blasted semicolons!
tax_table[] <- lapply(tax_table, gsub, pattern=';', replacement='')

zotu <- left_join(zotu, tax_table)



USearch pipeline to identify OTUs and assign taxonomy
```{r}


zotu <- read_csv("../processed_data/zotu.csv") 
tax <- read_csv("../processed_data/tax.csv") 
meta <- read_csv("../ref/metadata.csv")

# Check read counts
read.counts <- zotu %>%
  rowwise(seqid) %>%
  mutate(numreads = rowSums(across(where(is.numeric)))) %>%
  ungroup() %>%
  select(seqid, numreads)

#tested <- c("G0046", 'G1168', 'G1179', 'G0872', 'G1203', 'G0097', 'G0660', 'G0984', 'G0980', 'G1196', 'G0708', 'G1218', '1190')

# Pull out samples w/ >10k reads
samples.to.keep <- read.counts %>%
  filter(numreads > 1000) %>%
  select(seqid) %>%# filter(seqid %in% tested) %>%
  as_vector()


# Drop <10k read samples from zotu table
zotu <- zotu %>%
  filter(seqid %in% samples.to.keep)

# make tidy data
# Convert to relative abundance
zotu <- zotu %>%
  pivot_longer(!seqid, names_to="OTU", values_to="reads") %>%
  group_by(seqid) %>%
  mutate(relabd = (reads / sum(reads))) %>% ungroup


# calc log10 OTU abd
zotu <- zotu %>%
  mutate(log10relabd = if_else(relabd == 0, -6, log10(relabd)))

#filter out otus
otus.to.keep <- zotu %>%
  group_by(OTU) %>%
  summarise(avg_abd = mean(relabd)) %>%
  filter(avg_abd > .001) %>%
  select(OTU) %>%
  as_vector()

zotu <- zotu %>%
  filter(OTU %in% otus.to.keep)

# Join metadata to zotu table
zotu <- zotu %>%
  left_join(., meta, by= c("seqid"="sample")) %>% dplyr::rename("sample"="seqid")

#now add taxa
gtable <- tax %>%
  select(OTU, genus, family)

zotu <- zotu %>%
  left_join(gtable, by="OTU")
```


Heatmap of all samples
```{r}

library('ggdendro')
library('grid')


humanized <- zotu %>% select(OTU, sample, log10relabd) %>% 
              distinct() %>% reshape2::dcast(OTU~sample)

humanized.scaled <- humanized

#clustering
set.seed(123)
humanized.matrix <- as.matrix(humanized.scaled[, -c(1)])
rownames(humanized.matrix) <- humanized.scaled$OTU
humanized.dendro <- as.dendrogram(hclust(d = dist(x = humanized.matrix)))

# Create dendro
dendro.plot <- ggdendrogram(data = humanized.dendro, rotate = TRUE)
dendro.OTU <- dendro.plot

#reorder
humanized.order <- order.dendrogram(humanized.dendro)
zotu$OTU <- factor(x = zotu$OTU,
                               levels = humanized.scaled$OTU[humanized.order], 
                               ordered = TRUE)


#now reorder X axis
humanized <- zotu %>% select(OTU, sample, log10relabd) %>% 
              distinct() %>% reshape2::dcast(sample~OTU)
#clustering
set.seed(123)
humanized.matrix <- as.matrix(humanized[, -c(1)])
rownames(humanized.matrix) <- humanized$sample
humanized.dendro <- as.dendrogram(hclust(d = dist(x = humanized.matrix)))

# Create dendro
dendro.plot <- ggdendrogram(data = humanized.dendro, rotate = TRUE)
dendro.x <- dendro.plot

#reorder
humanized.order <- order.dendrogram(humanized.dendro)
zotu$sample <- factor(x = zotu$sample,
                               levels = humanized$sample[humanized.order], 
                               ordered = TRUE)

##GET BUG COLORS FROM BAR GRAPHS AT END
#bugColors <- setNames(c(scales::hue_pal()(length(unique(zotu$family)))), c(levels(as.factor(zotu$family))))

#fam lvl
heatmap.plot <- zotu %>% 
  ggplot(., aes(x=sample, y=OTU, fill=log10relabd)) +
  geom_tile(color = 'white') +
  scale_fill_gradient2(high = 'firebrick3', mid = 'white', low = '#1357a6', midpoint = -3) +
  labs(x = 'Human Donor', y = 'OTU', fill = 'log10(Relative Abundance)') +
  theme(strip.text.y.right = element_text(angle = 0, size = 10),
        strip.text.x.top = element_text(angle = 0, size = 10),
              axis.ticks = element_blank(), 
              axis.text.x = element_text(size = 10, hjust = 1, angle = 90),
              axis.text.y = element_blank(),
              strip.text = element_text(size = 6),
              panel.background = element_blank(), 
              strip.background = element_blank(),
              legend.position = 'bottom',
              panel.spacing = unit(0.05, "lines"))
heatmap.plot

label <- ggplot(data = zotu, aes(x = sample, y = '')) +
    geom_tile(aes(fill = group), color = 'white') + 
    scale_fill_manual(values = c('firebrick', 'dodgerblue', 'forestgreen')) +
    labs(fill = '') +
    theme(strip.text.y.right = element_text(angle = 0),
              panel.background = element_blank(), 
              axis.ticks = element_blank(), 
              axis.text.y = element_blank(),
              axis.title.y = element_blank(),
              axis.text.x = element_blank(), 
              axis.title.x = element_blank(),
              strip.text = element_blank(),
              panel.spacing = unit(0.05, "lines"))

family <-  zotu %>% mutate(family = ifelse(family %in% top10, family, 'Other')) %>%
    ggplot(., aes(x = '', y = OTU)) +
    geom_tile(aes(fill = family), color = 'white') + 
    scale_fill_manual(values = bugColors) +
    labs(fill = 'family') +
    theme(strip.text.y.right = element_text(angle = 0),
              panel.background = element_blank(), 
              axis.ticks = element_blank(), 
              axis.text.y = element_blank(),
              axis.title.y = element_blank(),
              axis.text.x = element_blank(), 
              axis.title.x = element_blank(),
              strip.text = element_blank(),
              panel.spacing = unit(0.05, "lines"))
family

dendro <- plot_grid(dendro.OTU, dendro.x)
dendro
ggsave("../figs/dendrogram.png",  width = 6 , height = 6)
ggsave("../figs/dendrogram.pdf",  width = 6, height = 6)

cowplot::plot_grid(heatmap.plot, family, label, ncol =3, rel_widths = c(.2,.3, .2))
ggsave("../figs/heatmap.png",  width = 14, height = 6)
ggsave("../figs/heatmap.pdf",  width = 14, height = 6)

```

PCA of samples
```{r}

library(M3C)

vendor <- zotu %>% 
  select(OTU, sample,group, relabd) %>% 
  reshape2::dcast(., group + sample ~ OTU) 

set.seed(123)


humanized.matrix <- as.matrix(humanized[, -c(1)])


#create such that samples as columns, rows as features
tSNE_header <- (vendor %>% select(-sample) %>% t() %>% as.data.frame())[1,]


tSNE_input <- (vendor %>% select(-sample) %>% t() %>% as.data.frame())[-1,] %>% 
              mutate_all(., function(x) as.numeric(as.character(x))) 

names(tSNE_input) <- tSNE_header[1,]


s9c <- M3C::pca(tSNE_input, 
         labels = as.factor(names(tSNE_input)),
         colvec = c('firebrick', 'dodgerblue', 'forestgreen'),
         dotsize = 3,
         controlscale = T, scale = 3, 
         axistextsize = 10,
         legendtextsize = 14, legendtitle = '')
plotly::ggplotly(s9c)

ggsave("../figs/PCA_Charlie.png", width = 4, height = 3)
ggsave("../figs/PCA_Charlie.pdf", width = 4, height = 3)

```

How is diversity?
```{r}

library(vegan)

tSNE_header <- (vendor %>% select(-group) %>% t() %>% as.data.frame())[1,]

sd_df<- tSNE_input %>% t() %>% vegan::diversity(index = "shannon")

sd_lab <- tSNE_header %>% t()
grp_key <- vendor[,1:2]

shannonD <- sd_df %>% as.data.frame() %>% cbind(., sd_lab) %>% select(shannD = '.', sample)

#normalize by log(k) where k is number of species over 0.05
numk <- zotu %>% 
  select(OTU, sample, relabd) %>% 
  melt(id.vars = c('sample', 'OTU')) %>% 
  filter(value > .001) %>% 
  select(-value) %>% distinct() %>% 
  group_by(sample) %>% mutate(kval = log2(n())) %>%
  ungroup() %>% select(sample, kval) %>% distinct() %>% na.omit()

shannonD <- left_join(shannonD, numk) %>% mutate(shann_entropy = shannD/kval) %>% left_join(.,grp_key)

#IF ONLY DOING SHANNON DIVERSITY
#shannonD <- left_join(shannonD, numk) %>% left_join(.,grp_key)


ggplot(shannonD, aes(factor(group, levels = c("H", "AR", "GC")), 
                     shann_entropy), fill = 'white') +
    stat_boxplot(geom= 'errorbar', width = .2, lwd = .5) +
    geom_boxplot(outlier.shape = NA, lwd = .5) + 
    geom_jitter(size = 2) +
  labs(y = 'Normalized Shannon Entropy', x = 'Disease Group') +
  theme_cowplot() +
  ggsignif::geom_signif(comparisons = list(c("AR","H"),c("H","GC"),c("AR","GC")), 
                         y_position = c(0.63, .605, .615),
                         map_signif_level=TRUE, #Comment out for p-values
                         test = "wilcox.test") +
  theme(axis.text.x = element_text(size = 14), 
  axis.text.y = element_text(size = 14),
axis.title.x.bottom = element_text(size = 16),
axis.title.y.left = element_text(size = 16),
legend.position = 'none')

ggsave("../figs/Entropy_Charlie.png", width = 2, height = 3.5)
ggsave("../figs/Entropy_Charlie.pdf", width = 2, height = 3.5)

```


Also do richness (alpha-diversity)
```{r}

chao <- zotu %>% 
  select(OTU, sample,group, reads) %>% 
  reshape2::dcast(., group + sample ~ OTU) 

set.seed(123)


#create such that samples as columns, rows as features
tSNE_header <- (chao %>% #select(-sample) %>% 
                  t() %>% as.data.frame())[1:2,] %>% t()


tSNE_input <- (chao %>% select(-sample) %>% t() %>% as.data.frame())[-1,] %>% 
              mutate_all(., function(x) as.numeric(as.character(x))) 


# calculate richness and Chao1 using vegan package
sd_df <- tSNE_input %>% t() %>% vegan::estimateR(.)

shannonD <- sd_df %>% as.data.frame() %>% t() %>% cbind(., tSNE_header) %>% as.data.frame()
shannonD[1:5] <- sapply(shannonD[1:5],as.numeric)

#wilcox.test(filter(shannonD, group == 'GC')$S.chao1, filter(shannonD, group == 'H')$S.chao1)

ggplot(shannonD, aes(factor(group, levels = c("H","AR", "GC")), 
                     S.chao1), fill = 'white') +
    stat_boxplot(geom= 'errorbar', width = .2, lwd = .5) +
    geom_boxplot(outlier.shape = NA, lwd = .5) + 
    geom_jitter(size = 2) +
  labs(y = 'Estimated Species Richness [chao1]', x = 'Disease Group') +
  theme_cowplot() +
  ggsignif::geom_signif(comparisons = list(c("AR","H"),c("H","GC"),c("AR","GC")), 
                         y_position = c(123, 125.5, 128),
    #                     map_signif_level=TRUE, #Comment out for p-values
                         test = "wilcox.test") +
  theme(axis.text.x = element_text(size = 14), 
  axis.text.y = element_text(size = 14),
axis.title.x.bottom = element_text(size = 16),
axis.title.y.left = element_text(size = 16),
legend.position = 'none')

ggsave("../figs/Richness_Charlie.png", width = 2, height = 3.5)
ggsave("../figs/Richness_Charlie.pdf", width = 2, height = 3.5)
```


Can we do DEseq to identify differentially abundant OTUs?


```{r}

library("DESeq2")

#Make count matrix
cts <- zotu %>% 
        select(sample, reads, OTU) %>% 
        dcast(OTU~sample, value.var = "reads") %>% 
        column_to_rownames("OTU")

#add pseudocounts
cts[is.na(cts)] <- 0
cts <-(cts + 1)

#Generate MetaData 
coldata <- cts %>% t() %>% as.data.frame %>% 
      rownames_to_column("sample") %>% select(sample) %>%
           left_join(., select(zotu, sample, group) %>% 
                  distinct) %>% 
       column_to_rownames("sample") 

#Make sure dfs are correct
all(rownames(coldata) %in% colnames(cts))
all(rownames(coldata) == colnames(cts))


```


```{r}
set.seed(123)
#Make DeSeq DataSet
dds <- DESeqDataSetFromMatrix(countData = cts,
                              colData = coldata,
                              design = ~ group)

#Differential 'expression' analysis
dds <- DESeq(dds)

#Build result table for C vs A
res_CvsA <- results(dds, contrast=c('group', "GC","AR"), 
                      pAdjustMethod = 'BH')

#Build result table for C vs B
res_CvsB <- results(dds, contrast=c('group', "GC","H"), 
                      pAdjustMethod = 'BH')

#Build result table for A vs B
res_AvsB <- results(dds, contrast=c('group', "AR","H"), 
                      pAdjustMethod = 'BH')

```


Let's see those plots!!!
```{r}


## Obtain logical vector where TRUE values denote padj values < 0.05 and fold change > 2 in either direction
res_table_CvsA <- res_CvsA %>% as.data.frame() %>%
                      mutate(threshold = padj < 0.05 & abs(log2FoldChange) >= 1,
                             OTUType = ifelse(log2FoldChange > 0 & threshold == T, 
                                           "GC-enriched", NA),
                             OTUType = ifelse(log2FoldChange < 0 & threshold == T,
                                           "AR-enriched", OTUType)) %>% 
                  tibble::rownames_to_column(., "OTU") %>% 
                inner_join(., tax)

res_table_CvsB <- res_CvsB %>% as.data.frame() %>%
                      mutate(threshold = padj < 0.05 & abs(log2FoldChange) >= 1,
                             OTUType = ifelse(log2FoldChange > 0 & threshold == T, 
                                           "GC-enriched", NA),
                             OTUType = ifelse(log2FoldChange < 0 & threshold == T,
                                           "H-enriched", OTUType)) %>% 
                  tibble::rownames_to_column(., "OTU") %>% 
                inner_join(., tax)

res_table_AvsB <- res_AvsB %>% as.data.frame() %>%
                      mutate(threshold = padj < 0.05 & abs(log2FoldChange) >= 1,
                             OTUType = ifelse(log2FoldChange > 0 & threshold == T, 
                                           "AR-enriched", NA),
                             OTUType = ifelse(log2FoldChange < 0 & threshold == T,
                                           "H-enriched", OTUType)) %>% 
                  tibble::rownames_to_column(., "OTU") %>% 
                inner_join(., tax)
                  
## Volcano plot 
volcano_A <- ggplot(res_table_CvsA) +
    geom_point(aes(x = log2FoldChange, y = -log10(padj), color = OTUType),size = 2) +
    #scale_color_manual(values = c('#6BBCD1','#FD9415' )) +
    scale_color_manual(values = c('firebrick', 'dodgerblue')) +
    ggtitle("AR vs GC") +
    xlab("log2(Fold Change)") + 
    ylab("-log10(adjusted p-value)") +
    xlim(-12,12) +
    ylim(0,10) +
  theme_cowplot() +
    theme(legend.position = "none",
          plot.title = element_text(size = rel(1.5), hjust = 0.5),
          axis.title = element_text(size = rel(1.25)))          

plotly::ggplotly(volcano_A)

volcano_B <- ggplot(res_table_CvsB) +
    geom_point(aes(x = log2FoldChange, y = -log10(padj), color = OTUType),size = 2) +
    #scale_color_manual(values = c('#6BBCD1','#FD9415' )) +
    scale_color_manual(values = c('dodgerblue', 'forestgreen')) +
    ggtitle("H vs GC") +
    xlab("log2(Fold Change)") + 
    ylab("-log10(adjusted p-value)") +
    xlim(-11,11) +
    ylim(0,10) +
  theme_cowplot() +
    theme(legend.position = "none",
          plot.title = element_text(size = rel(1.5), hjust = 0.5),
          axis.title = element_text(size = rel(1.25)))          

plotly::ggplotly(volcano_B)


volcano_c <- ggplot(res_table_AvsB) +
    geom_point(aes(x = log2FoldChange, y = -log10(padj), color = OTUType),size = 2) +
    #scale_color_manual(values = c('#6BBCD1','#FD9415' )) +
    scale_color_manual(values = c('firebrick', 'forestgreen')) +
    ggtitle("H vs AR") +
    xlab("log2(Fold Change)") + 
    ylab("-log10(adjusted p-value)") +
    xlim(-11,11) +
    ylim(0,10) +
    theme_cowplot() +
    theme(legend.position = "none",
          plot.title = element_text(size = rel(1.5), hjust = 0.5),
          axis.title = element_text(size = rel(1.25)))   


plotly::ggplotly(volcano_c)


plot_grid(volcano_A, volcano_B, volcano_c, nrow = 1)

ggsave("../figs/DEseq_Charlie.png", width = 10.5, height = 3.5)
ggsave("../figs/DEseq_Charlie.pdf", width = 10.5, height = 3.5)

```

```{r}


fam_abd <- vendor %>% reshape2::melt() %>% group_by(group, variable) %>% 
    mutate(median_relabd = median(value)) %>% select(-value, -sample) %>% ungroup %>% 
  distinct() %>%
  left_join(., tax, by = c('variable'='OTU')) %>%
  select(family, group, median_relabd) %>% group_by(group, family) %>% 
  mutate(ttl_fam = sum(median_relabd)) %>%
  filter(ttl_fam > .0001) %>% select(-median_relabd) %>% ungroup %>% distinct() %>% 
  mutate(family = ifelse(ttl_fam < .003, 'Other', family)) %>% group_by(family, group) %>% mutate(ttl_fam = sum(ttl_fam)) %>% ungroup() %>% distinct()

set.seed(123)

#Get top 10 families
top10 <- c((fam_abd %>% group_by(family) %>% mutate(famcount=sum(ttl_fam)) %>% 
  select(family, famcount) %>% distinct() %>% arrange(desc(famcount)))$family[1:10], 'Enterobacteriaceae', 'Bifidobacteriaceae')

bugColors <- setNames(c(scales::hue_pal()(length(unique(top10)))), c(levels(
  (sample(x = as.factor(top10),size = length(top10),replace = F))
  )))

bugColors <- bugColors[top10]


plot <- fam_abd %>% mutate(family = ifelse(family %in% top10, family, 'Other')) %>% #only look at top 10 families
  group_by(family, group) %>%
  mutate(ttl_fam = sum(ttl_fam)) %>% ungroup %>% distinct() %>%
  ggplot(., aes(x = factor(group, levels = c('H', 'AR', 'GC')), y=ttl_fam, fill = family),color = 'black') + 
  geom_col(stat="identity", position = 'fill', width = .5) +
  labs(x='', y = 'Rel. Abundance', fill = 'Family') +
    scale_fill_manual(values = bugColors) +
  theme_cowplot() +
    theme(plot.background = element_blank(),
          strip.background = element_blank(),
          axis.title.y = element_text(size = 14),
              axis.text.y = element_text(size = 12), 
              axis.text.x = element_text(size = 12),
              axis.ticks.x = element_blank(),
              panel.spacing = unit(0.05, "lines"),
          legend.text = element_text(size = 8),
          legend.key.height  = unit(.35, 'cm'),
          legend.key.width  = unit(.35, 'cm'))

plotly::ggplotly(plot)
#plot
ggsave("../figs/16Scomposition.png", width = 3.5, height = 2.5)
ggsave("../figs/16Scomposition.pdf", width = 3.5, height = 2.5)


```

Streptococcae and erysipelotrichacea seem enriched, is that legit?
```{r}
#Cool, streptococcae foound in A but not B... Why dat?

#sig_OTUs <- (compare %>% filter(threshold == T)) %>% select(OTU)

plota <- zotu %>% 
  #filter(family == 'Streptococcaceae') %>% 
  #filter(family == 'Ruminococcaceae') %>% 
  filter(OTU == 'Otu9') %>%
  #group_by(sample, OTU) %>% 
  #mutate(ttl_strep = sum(relabd)) %>% ungroup() %>% 
  select(sample, group, gender, relabd, OTU) %>% 
  distinct() %>%
  ggplot(., aes(group, relabd)) +
  scale_y_log10() +
  stat_boxplot(width = .1, geom ='errorbar') +
  geom_boxplot(outlier.shape = NA, width = .5) + 
  geom_jitter(aes(color = gender)) + 
  ggsignif::geom_signif(comparisons = list(c("AR", "H"),c("AR","GC"),c("H","GC")), 
                         y_position = c(-.5, -.2, -.8),
                         map_signif_level=TRUE, #Comment out for p-valu es
                         test = "wilcox.test") + 
  theme_cowplot() +
  labs(x='Group', y='Rel. Abd.', title = 'OTU9') +
  theme(axis.text = element_text(size = 12),
        axis.title = element_text(size = 14), strip.background = element_blank()) 

plota

ggsave("../figs/OTU9_Enrichment.pdf", width = 3.5, height = 3.5)

#plotly::ggplotly(plota)
#lefse analysis
```

```{r}
fam <- 'Streptococcaceae'

plota <- zotu %>% 
  filter(family == fam) %>% 
  group_by(sample) %>% 
  mutate(ttl_strep = log10(mean(relabd))) %>% ungroup() %>% 
  select(group, ttl_strep) %>% 
  distinct() %>%
  ggplot(., aes(factor(group, levels = c("H", "AR", "GC")), ttl_strep)) +
  stat_boxplot(width = .1, geom ='errorbar') +
  geom_boxplot(outlier.shape = NA, width = .5) + 
  geom_jitter() + 
  ggsignif::geom_signif(comparisons = list(c("AR", "H"),c("AR","GC"),c("H","GC")), 
                         y_position = c(-1.2, -1, -.75),
                         map_signif_level=TRUE, #Comment out for p-valu es
                         test = "wilcox.test") + 
  theme_cowplot() +
  labs(x='Cohort', y='log10(Rel. Abd.)', title = fam) +
  theme(axis.text = element_text(size = 14),
        axis.title = element_text(size = 14), strip.background = element_blank())
plota 

fam <- 'Bifidobacteriaceae'

plotb <- zotu %>% 
  filter(family == fam) %>% 
  group_by(sample) %>% 
  mutate(ttl_strep = log10(mean(relabd))) %>% ungroup() %>% 
  select(group, ttl_strep) %>% 
  distinct() %>%
  ggplot(., aes(factor(group, levels = c("H", "AR", "GC")), ttl_strep)) +
  stat_boxplot(width = .1, geom ='errorbar') +
  geom_boxplot(outlier.shape = NA, width = .5) + 
  geom_jitter() + 
  ggsignif::geom_signif(comparisons = list(c("AR", "H"),c("AR","GC"),c("H","GC")), 
                         y_position = c(-1.2, -1, -.75),
                         map_signif_level=TRUE, #Comment out for p-valu es
                         test = "wilcox.test") + 
  theme_cowplot() +
  labs(x='Cohort', y='log10(Rel. Abd.)', title = fam) +
  theme(axis.text = element_text(size = 14),
        axis.title = element_text(size = 14), strip.background = element_blank())

plotb

fam <- 'Enterobacteriaceae'

plotc <- zotu %>% 
  filter(family == fam) %>% 
  group_by(sample) %>% 
  mutate(ttl_strep = log10(mean(relabd))) %>% ungroup() %>% 
  select(group, ttl_strep) %>% 
  distinct() %>%
  ggplot(., aes(factor(group, levels = c("H", "AR", "GC")), ttl_strep)) +
  stat_boxplot(width = .1, geom ='errorbar') +
  geom_boxplot(outlier.shape = NA, width = .5) + 
  geom_jitter() + 
  ggsignif::geom_signif(comparisons = list(c("AR", "H"),c("AR","GC"),c("H","GC")), 
                         y_position = c(-1.2, -1, -.75),
                         map_signif_level=TRUE, #Comment out for p-valu es
                         test = "wilcox.test") + 
  theme_cowplot() +
  labs(x='Cohort', y='log10(Rel. Abd.)', title = fam) +
  theme(axis.text = element_text(size = 14),
        axis.title = element_text(size = 14), strip.background = element_blank())

plotc

plot_grid(plota, plotb, plotc, rel_widths = c(1,1,1), nrow = 1)

ggsave("../figs/FamilyEnrichments.png", width = 9, height = 4)
ggsave("../figs/FamilyEnrichments.pdf", width = 9, height = 4)

plota
ggsave("../figs/StrepEnrichments.png", width = 2.5, height = 4)
ggsave("../figs/StrepEnrichments.pdf", width = 2.5, height = 4)

```

Can also look at phylum, but not really that different
```{}
phy_abd <- vendor %>% reshape2::melt() %>% group_by(group, variable) %>% 
    mutate(median_relabd = median(value)) %>% select(-value, -sample) %>% ungroup %>% 
  distinct() %>%
  left_join(., tax_table, by = c('variable'='OTU')) %>%
  select(Phylum, group, median_relabd) %>% group_by(group, Phylum) %>% mutate(ttl_fam = sum(median_relabd)) %>% select(-median_relabd) %>% ungroup %>% distinct()






plotb <- ggplot(phy_abd, aes(x = group, y=ttl_fam, fill = Phylum)) + 
  geom_col(stat="identity", position = 'fill', width = .5) +
  #geom_col(stat="identity", position = 'stack') +
  labs(x='', y = 'Rel. Abundance', fill = 'Phylum') +
  #  scale_fill_manual(values = bugColors) +
  theme_cowplot() +
    theme(plot.background = element_blank(),
          strip.background = element_blank(),
          axis.title.y = element_text(size = 14),
              axis.text.y = element_text(size = 12), 
              axis.text.x = element_text(size = 12, angle = 45, hjust = 1),
              axis.ticks.x = element_blank(),
              panel.spacing = unit(0.05, "lines"),
          legend.text = element_text(size = 8),
          legend.key.height  = unit(.35, 'cm'),
          legend.key.width  = unit(.35, 'cm'))

cowplot::plot_grid(plot,  plotb)

```

